\plaintitle

### 2016-09-21 Paper/proceedings.tex
Attempt at a "see, people are pretty good at this" paragraph.

\usepackage{amsmath}
% If you want to use todo notes, marginpars etc. during creation of your draft document, you
% have to enable the "chi_draft" option for the document class. To do this, change the very first
% line to: "\documentclass[chi_draft]{sigchi}". You can then place todo notes by using the "\todo{...}"
% command. Make sure to disable the draft option again before submitting your final document.
\usepackage{todonotes}

@@

Recent work in the visualization community has focused on the perception of correlation in scatter plots. Rensink et al.~\cite{rensink2010perception} show that viewers can estimate correlation with some accuracy in scatter plots. Harrison et al.~\cite{harrison2014ranking} extend this finding to other visualization types, and a re-analysis by Kay \& Heer~\cite{kay2016beyond} indicates that, for many of the more esoteric bivariate visualizations, performance at this task is poor. Estimation of correlation can also be biased by the choice of axis scales~\cite{cleveland1982variables}: the whitespace and aspect ratio changes introduced by expanding the scale of the axes can cause over-estimation of correlation.

@@

For each point set, we created a set of residual values, sampled evenly from a Gaussian. The bandwidth of this Gaussian controls both goodness of fit and dispersion of points (see Fig. \ref{fig:sigmas}). We permuted this set of residuals, and then applied them to points along a target trend. As heteroskedasticity introduced through permutation could alter the actual trend away from the target trend, we performed rejection sampling to ensure that the trend of the resulting points were within $10^{-7}$ of the target trend. We reused these residuals across all different trend types (linear, quadratic, or trigonometric). Fig. \ref{fig:trendtypes} shows these different trend types on the same set of residuals.

@@

Our results support our first hypothesis: \textbf{larger residuals reduce accuracy at regression by eye}.  We observed a significant main effect for the bandwidth of the Gaussian used to generate residuals ($F(1,4554)=950$, $p<0.001$). The interquartile mean of absolute error increased monotonically with this bandwidth, from $7\%$ of the actual slope of the trend line when the bandwidth was $0.05$, to $32\%$ when the bandwidth was $0.20$. Figure \ref{fig:exp1sigmas} illustrates this result.

Our results fail to support our second hypothesis: \textbf{there was no statistically significant difference in estimation accuracy among linear, quadratic, or trigonometric trends}. Fit type had only a marginal main effect ($F(2,4554)=2.60$, $p=0.074$), and post-hoc tests (using Tukey's Honest Significant Difference) did not identify any significant pairwise interactions. The interquartile mean of absolute error of linear fits was $17\%$ of the actual slope, in between a $16\%$ error for quadratic fits, and $18\%$ for trignometric fits (see Fig. \ref{fig:exp1types}). This result suggests that the relative unfamiliarity of non-linear trends does not have a notable impact on performance.

@@

In general, the absolute error of slope estimation was low, with an interquartile mean of $0.06$ for slope values ranging from $\{-1,1\}$, corresponding to an error rate of $17\%$ of the actual slope.

@@

This experiment was largely identical to Experiment 1, except we designated 0, 5, 10, or 15 points at the very beginning, first third, or end of the series as outliers. We vertically positioned outliers within the top or bottom 10\% of the visual area of the visualization (whichever was farthest from the trend line), and added random jitter. We then calculated the outlier-sensitive line of best fit, as well as the intersection between this new trend line and the original, outlier-less trend line. Figure \ref{fig:outliers} shows examples of this process.

@@

Similar to the previous experiments, participants saw one of each combination of factors: the three graph types, eight slopes, and four Gaussian residual bandwidths. The four outlier quantities $\{0,5,10,15\}$ was an additional factor. To maintain a manageable number of stimuli (piloting showed evidence of fatigue for more than 100 stimuli per trial), both the sign of the trend line (positive or negative), and the location of the outliers (beginning, first third, or end) were random factors, with each level apportioned to half and one third of the stimuli respectively. This resulted in 96 total stimuli, with an additional 4 validation stimuli otherwise excluded from analysis, in line with prior experiments.

@@

Our experimental setup was intentionally simple, affording only a single free parameter for each experiment. In actual regression by eye, the viewer may simultaneously engage in multiple types of estimation: choosing a particular type of fit, ignoring outliers, and estimating the parameters of the chosen model. Errors in any one of these steps could compound, resulting in performance worse than our measures, where many of these decisions are fixed.

>>>>

%MC Packages
\usepackage{amsmath}
\usepackage{cite}

% If you want to use todo notes, marginpars etc. during creation of your draft document, you
% have to enable the "chi_draft" option for the document class. To do this, change the very first
% line to: "\documentclass[chi_draft]{sigchi}". You can then place todo notes by using the "\todo{...}"
% command. Make sure to disable the draft option again before submitting your final document.
\usepackage{todonotes}

@@

Recent work in the visualization community has focused on the perception of correlation in scatter plots. Rensink et al.~\cite{rensink2010perception} show that viewers can estimate correlation with some accuracy in scatter plots. Harrison et al.~\cite{harrison2014ranking} extend this finding to other visualization types, and a re-analysis by Kay \& Heer~\cite{kay2016beyond} indicates that, for many of the more esoteric bivariate visualizations, performance at this task is poor. Estimation of correlation can also be biased by the choice of axis scales~\cite{cleveland1982variables}: the whitespace changes introduced by expanding the scale of the axes can cause over-estimation of correlation.

@@

For each point set, we placed 100 points along a particular trend, regularly spaced in x. This both affords the use of linegraphs and area charts for displaying this data (as they require that points be one to one), and aligns well with time series data, a common domain for estimating (temporal) trends and relationships. For each point set, we created a set of residual values, sampled evenly from a Gaussian. The bandwidth of this Gaussian controls both goodness of fit and dispersion of points (see Fig. \ref{fig:sigmas}). We permuted this set of residuals, and then applied them to the original points. As heteroskedasticity introduced through permutation could alter the actual trend away from the target trend, we performed rejection sampling to ensure that the trend of the resulting points were within $10^{-7}$ of the target trend. We reused these residuals across all different trend types (linear, quadratic, or trigonometric). Fig. \ref{fig:trendtypes} shows these different trend types on the same set of residuals.

@@

Our results support our first hypothesis: \textbf{larger residuals reduce accuracy at regression by eye}.  We observed a significant main effect for the bandwidth of the Gaussian used to generate residuals ($F(1,4554)=950$, $p<0.001$). The interquartile mean of absolute error increased monotonically with this bandwidth, from $0.02$ of the actual slope of the trend line when the bandwidth was $0.05$, to $0.12$ when the bandwidth was $0.20$. Figure \ref{fig:exp1sigmas} illustrates this result.

Our results fail to support our second hypothesis: \textbf{there was no statistically significant difference in estimation accuracy among linear, quadratic, or trigonometric trends}. Fit type had only a marginal main effect ($F(2,4554)=2.60$, $p=0.074$), and post-hoc tests (using Tukey's Honest Significant Difference) did not identify any significant pairwise interactions. The interquartile mean of absolute error of was $0.06$ for linear and quadratic fits, and $0.07$ for trigonometric fits (see Fig. \ref{fig:exp1types}). This result suggests that the relative unfamiliarity of non-linear trends does not have a notable impact on performance.

@@

%The average absolute error of slope estimation across all conditions had an interquartile mean of $0.06$ for slope values ranging from $\{-1,1\}$, corresponding to an error rate of $17\%$ of the actual slopes. 
As a point of comparison, the interquartile mean of the absolute log error, calculated as per Cleveland \& McGill~\cite{cleveland1984graphical} ($log_2 |\text{error}| + \frac{1}{8}$) was $2.4$ across all conditions. This is comparable to the error rates of many other graphical perception tasks as observed in both lab and crowdsourced studies. For instance, Heer \& Bostock~\cite{heer2010crowdsourcing}  report an absolute log error of $2.5$ for proportional judgments in treemaps, and Cleveland \& McGill~\cite{cleveland1984graphical} report an absolute log error of $2.4$ for proportional judgments in stacked bars (although compare to a log error of estimation of relative lengths of lines of $1.1$). This similarity suggests that, despite requiring estimation of aggregate statistical information, regression by eye results in judgments that are accurate enough for many practical purposes, comparable with the expected accuracy in comparing individual values in visualizations.

@@

This experiment was largely identical to Experiment 1, except we designated 0, 5, 10, or 15 points at the very beginning, first third, or end of the series as outliers. We vertically positioned outliers within the top or bottom 10\% of the visual area of the visualization (whichever was farthest from the trend line), and added random jitter. We then calculated the outlier-sensitive line of best fit, as well as the intersection between this new trend line and the original, outlier-less trend line. We refer to the original trend line, estimated without the presence of outliers, as the \textbf{robust trend line}. We refer to the re-estimated trend, which takes into account the added outliers, as the \textbf{OLS trend line}. Figure \ref{fig:outliers} shows examples of this process, and the corresponding changes in trend line.

@@

Participants saw one of each combination of factors: the three graph types, eight slopes, and four Gaussian residual bandwidths. The four outlier quantities $\{0,5,10,15\}$ were an additional factor. To maintain a manageable number of stimuli (piloting showed evidence of fatigue for more than 100 stimuli per trial), both the sign of the trend line (positive or negative), and the location of the outliers (beginning, first third, or end) were random factors, with each level randomly apportioned to half and one third of the stimuli respectively. This resulted in 96 total stimuli, with an additional 4 validation stimuli otherwise excluded from analysis, in line with prior experiments.

@@

Our experimental setup was intentionally simple, affording only a single free parameter for each experiment. In actual regression by eye, the viewer may simultaneously engage in multiple types of estimation: choosing a particular type of fit, ignoring outliers, and estimating the parameters of the chosen model. Errors in any one of these steps could compound, resulting in performance worse than our measures, where many of these decisions are fixed. It is also likely that there are many visual features, such as the convex hull of points, or the numerosity of point clouds, that are acting as visual proxies for estimation of the trend. Given the unimodal error function and uniform density of our experimental stimuli, these proxies are useful for estimating the trend line. However, in datasets where internal densities of points may be skewed, or the envelope of point clouds non-informative for estimating central tendency, these visual proxies may introduce biases and confounds in visual estimations.
